\textbf{Resilient Machine Learning Approaches \\ for Fast Risk Evaluation and Management \\ of Financial Portfolios and Variable Annuities}

### 2025-03-24 pre_private/pre.tex
new

\jointwork{Supervised by Prof. Tony Wirjanto and Prof. Mingbin Feng}
\conference{Thesis Defense, \\ University of Waterloo}

@@

% Title page, navigation surpressed, no page number
{
\beamertemplatenavigationsymbolsempty
\begin{frame}[plain]
\titlepage
\end{frame}
}

{
\beamertemplatenavigationsymbolsempty
\defbeamertemplate*{headline}{miniframes theme no subsection no content}
{ \begin{beamercolorbox}{section in head/foot}
    \vskip\headheight
  \end{beamercolorbox}}
\begin{frame}{Outline}
\tableofcontents
\end{frame}
}
\addtocounter{framenumber}{-2}

@@

Involves two levels of Monte Carlo simulations:
\begin{itemize}
    \item Outer: underlying risk factors (outer scenarios), $X_i \sim F_X$
    \item Inner: scenario-wise losses (inner replications), $Y_{ij} \sim F_{Y|X_i}$
    \item Total simulation budget: $\Gamma = M \cdot N$
\end{itemize}

\vspace{10pt}
\textbf{Computationally expensive}

\note{In the context of quantitative risk management, nested simulation is used for risk evaluation of financial derivatives and insurance products. \\}
\note{The nested structure is due to the fact that the loss random variable is in the form of a conditional expectation of the portfolio loss given the underlying risk factors. \\}
\note{In our problems, the loss random variable $L$ depends on the risk factors $X$ and is the conditional expectation of the portfolio loss given $X$. \\}
\note{More specifically, $X$ is $d$-dimensional and $L$ is 1-dimensional. \\}
\note{Estimating a risk measure, $\rho(L)$, involves two levels of Monte Carlo simulations: outer level generates underlying risk factors, and inner level generates scenario-wise samples of portfolio losses. \\}
\note{The nested structure makes the estimation computationally expensive. \\}
\note{In this thesis, we focus on examining and improving the efficiency of nested simulation procedures mostly using supervised learning metamodels.}

@@

\begin{frame}{Common Risk Measures}

\begin{itemize}
    \item Smooth $h$, e.g., quadratic tracking error,

    $$ \rho(L) = \mathbb{E} \left[ (L - b)^2 \right]; $$

    \item hockey-stick $h$: mean excess loss,

    $$ \rho(L) = \mathbb{E} \left[ L \cdot \mathbbm{1}_{\{L \geq u\}} \right]; $$

    \item indicator $h$: probability of large loss,

    $$ \rho(L) = \mathbb{E} \left[ \mathbbm{1}_{\{L \geq u\}} \right]; $$

    \item Value at Risk (VaR),

    $$ \rho_\alpha(L) = Q_\alpha(L) = \inf \{ u: \mathbb{P}(L \leq u) \geq \alpha \}; $$

    \item Conditional Value at Risk (CVaR) \footnotemark,

    $$ \rho_\alpha(L) = \mathbb{E} \left[ L \vert L \geq Q_\alpha(L) \right]. $$

    \end{itemize}

    \footnotetext{Note: If $Q_\alpha(L)$ falls in a probability mass, $\rho(L) = \frac{(\beta - \alpha)Q_\alpha(L) + (1- \beta) \mathbb{E} \left[ L \vert L \geq Q_\alpha(L) \right]}{1-\alpha}$.}

\note{We consider different types of risk measures in this thesis. \\}
\note{The first three are in expectation form, and the last two are quantile-based risk measures. \\}
\note{The difference between the first three is the choice of the function $h$. \\}
\note{$h$ can be a smooth function, a hockey-stick function, or an indicator function. \\}
\note{The examples of risk measures are quadratic tracking error, mean excess loss, and probability of large loss. \\}
\note{The last two are value at risk and conditional value at risk, which are widely used in practice. \\}
\note{Note that the definition of CVaR requires the knowledge of the Value at Risk.}

@@

\begin{frame}{Standard Nested Simulation}

$$\hat{L}_{N, i} = \frac{1}{N} \sum_{j=1}^N Y_{ij}; ~~~ Y_{ij} \sim F_{Y|X_i} $$

\begin{itemize}
    \item Uses inner sample mean to estimate $L(X_i)$.
    \item Proposed by \citet{gordy2010nested}; finds optimal growth order of $M$ and $N$.
    \item \citet{zhang2021bootstrap} estimate the optimal $M$ and $N$ using a bootstrap method.
    \item Computationally expensive and potentially \textbf{wasteful} use of budget.
\end{itemize}

\note{In our first project, we focus on comparing existing nested simulation procedures for different types of risk measures. \\}
\note{They differ in the way they estimate $L(X_i)$, the portfolio loss conditional on the risk factors $X_i$. \\}
\note{The standard nested simulation procedure uses the inner sample mean under that scenarioto estimate $L(X_i)$. \\}
\note{The optimal growth order of $M$ and $N$ is found by \citet{gordy2010nested}. \\}
\note{For a growing total simulation budget $\Gamma$, $M$ should grow in the order of $\Gamma^{2/3}$, and $N$ should grow in the order of $\Gamma^{1/3}$. \\}
\note{For a finite-sample case, the bootstrap method is used to estimate the optimal $M$ and $N$ by \citet{zhang2021bootstrap}. \\}
\note{Since the standard procedure only uses the information from the same scenario, it is potentially \textbf{wasteful} use of budget. \\}
\note{Subsequent works focus on improving the efficiency of nested simulation by pooling inner replications from different outer scenarios.}

@@

\note{A supervised learning based nested simulation procedure is implemented following these steps. \\}
\note{First, the user runs the standard nested simulation procedure to generate a set of outer scenarios and inner replications. \\}
\note{Then, the outer scenarios and inner sample means can be treated as features and labels. \\}
\note{Their relationship is then approximated by a supervised learning metamodel. \\}
\note{Finally, the trained metamodel is used to make predictions for all $X \in \mathcal{X}$. \\}
\note{In other words, instead of using the inner sample mean to estimate $L(X_i)$, we use the trained metamodel to make predictions of the portfolio loss. \\}
\note{These loss predictions are then used to calculate the risk measures of interest. \\}
\note{For most research in this area, the focus is on the convergence rate of the risk measure estimator in terms of the simulation budget. \\}
\note{However, in this thesis, we note that there are other important aspects to consider, such as the cost of pooling that may come from metamodel training and validation. }

@@

\note{We start by comparing the asymptotic convergence rates of different nested simulation procedures. \\}
\note{These are the convergence rates of the MSE of the risk measure estimator when the total simulation budget $\Gamma$ goes to infinity, and when one follows the optimal growth order of $M$ and $N$. \\}
\note{The standard procedure has the same convergence rate of $\mathcal{O}(\Gamma^{-2/3})$ for all types of $h$ considered. The rates in red are not available in the original works but are derived in this thesis. \\}
\note{The regression-based procedure has $\mathcal{O}(\Gamma^{-1})$ convergence rate for smooth and hockey-stick $h$. \\}
\note{It is achieved by only allowing $M$ to grow in the order of $\Gamma^{1}$. In their experiments, $N$ is set to equal a value of $1$. \\}
\note{The kernel smoothing procedure is the only procedure that depends on the asset dimension $d$. \\}
\note{When risk factor $X$ is high-dimensional, the kernel smoothing procedure may converge slower than the standard procedure. \\}
\note{The kernel ridge regression procedure and likelihood ratio procedure have the same convergence rate as the regression-based procedure. \\}
\note{However, as we will see in the numerical experiments, their additional cost of pooling inner replications may offset their benefit of fast convergence rate. \\}
\note{Additionally for KRR, the authors analyze convergence of absolute error in probabilistic order instead of MSE.}

@@

\note{In most theoretical works, the focus is on the convergence rate of the MSE of the risk measure estimator. \\}
\note{However, for KRR, the authors analyze convergence of absolute error in probabilistic order. \\}
\note{One of the contributions of this thesis is to bridge the gap between MSE and absolute error convergence. \\}
\note{Let's first define the convergence in MSE and the convergence in probabilistic order. \\}
\note{The convergence in MSE is the convergence of the MSE of the risk measure estimator to the true risk measure, that is, $\hat{\rho}_{\Gamma}$ minus $\rho$ squared. \\}
\note{The convergence in probabilistic order is the convergence of the absolute error of the risk measure estimator to the true risk measure, that is, $|\hat{\rho}_{\Gamma} - \rho|$ to 0 in probability. \\}
\note{We show that the convergence rate of the MSE of the risk measure estimator implies the convergence rate of the absolute error in probabilistic order. \\}
\note{We will also show that the converse is not necessarily true.}

@@

\begin{frame}{Experiment Design}

We compare 5 nested simulation procedures 
\begin{itemize}
    \item Standard nested simulation
    \item Regression-based
    \item Kernel smoothing
    \item Likelihood ratio
    \item Kernel ridge regression
\end{itemize}

And their empirical convergence across different:
\begin{itemize}
    \item Risk measures
    \item Option types
    \item Asset dimensions
    \item Asset models (GBM vs. Heston)
    \item Regression bases (only for the regression-based procedure)
\end{itemize}

\note{Previously, we worked on the asymptotic convergence rates of different nested simulation procedures. \\}
\note{In practice, the total simulation budget is finite.}
\note{When the budget is finite, the theoretical convergence rate may not be the only thing we should consider. \\}
\note{We should also consider the cost of pooling itself. \\}
\note{The additional cost of pooling inner replications is not considered in the theoretical analysis of the previous works. \\}
\note{Neither it is mentioned in their numerical experiments. \\}
\note{More specifically, we have found that different research uses different numerical experiments to show the empirical rate of different procedures. \\}
\note{We aim to provide a fair and comprehensive analysis of the finite-sample performance of different nested simulation procedures. \\}
\note{The following numerical experiments test the empirical convergence rates of different nested simulation procedures across different risk measures, option types, asset dimensions, and asset models.}

@@

\begin{figure}
    \includegraphics[width=\textwidth]{../project1/figures/figure1.png}
    \caption{Empirical convergence rates of different procedures for the base case}
\end{figure}

\note{This figure shows the performance of all 5 procedures in their estimation of the quadratic tracking error for the portfolio of European call options.}
\note{It is a log-log plot of the MSE of the estimator versus the total simulation budget $\Gamma$.}
\note{For each procedure, we run 1000 macro replications of Monte Carlo simulations with different total simulation budgets. \\}
\note{For each data point on the plot, 1000 independent replications of the procedure are used to estimate the MSE. \\}
\note{The slopes of the lines estimates the empirical convergence rates of the procedures. \\}
\note{In the legend records the slopes of the lines and the corresponding asymptotic convergence rates of the procedures. \\}
\note{In the most basic case, most procedures have their empirical convergence rates match their asymptotic convergence rates. \\}
\note{However, as we will see after, this is not the case for all experiments. \\}
\note{The following slides will show the sensitivity of the procedures to different factors. \\}

@@

% \begin{frame}{Sensitivity to Asset Dimension}

% \begin{figure}
%     \includegraphics[width=\textwidth]{../project1/figures/figure2a.png}
% \end{figure}

% \begin{itemize}
%     \item Standard, KRR, and likelihood ratio procedures are dimension-independent
%     \item Kernel smoothing and regression show sensitivity to dimension, but in different ways
% \end{itemize}

% \note{This figure shows the sensitivity of the procedures to the asset dimension. \\}
% \note{The asset dimension is the dimension of the risk factors $X$. \\}
% \note{Everything is the same as the previous figure, except that the asset dimension is varied from 1 to 2. \\}
% \note{The standard, regression, KRR, and likelihood ratio procedures are dimension-independent. \\}
% \note{The kernel smoothing procedure is sensitive to the asset dimension. \\}
% \note{Despite the fact that the kernel smoothing procedure has the same asymptotic convergence rate as before, it converges slower in practice. \\}

@@

    \begin{figure}
        \includegraphics[width=\textwidth]{../project1/figures/figure2b.png}
    \end{figure}

\note{This figure shows the same experiment as the previous figure, but we further increase the asset dimension to 20. \\}
\note{The standard procedure is not affected by the asset dimension. \\}
\note{The kernel smoothing procedure actually converges faster than what its asymptotic convergence rate suggests. \\}
\note{The regression-based procedure is the only procedure that actually converges faster when the asset dimension is high. We will do a deeper dive into it in the following experiment. \\}
\note{It is also worth noting that the likelihood ratio and KRR procedures are dimension-independent. \\}
\note{While it is not shown in their empirical rate, due to their additional cost of pooling inner replications, we are not able to run these procedures when $\Gamma$ is large. \\}

\begin{frame}{Fast Convergence of Regression-based Procedure}

\begin{figure}
    \includegraphics[width=\textwidth]{../project1/figures/figure3.png}
\end{figure}

\begin{itemize}
    \item Higher initial convergence rate
    \item Stabilizes to match asymptotic rate at higher budgets
    \item Consistent across different asset dimensions
\end{itemize}

\note{To investigate the regression-based procedure, we further increase the total simulation budget $\Gamma$. \\}
\note{The figure shows the performance of the regression-based procedure in the estimation of the quadratic tracking error for the portfolio of European call options. \\}
\note{We separate different values of $\Gamma$ into two different parts. \\}
\note{The first part is when $\Gamma$ is small. \\}
\note{The second part is when $\Gamma$ is large. \\}
\note{The first part is when the regression-based procedure has a higher convergence rate than its asymptotic convergence rate. \\}
\note{The second part is when the regression-based procedure stabilizes and matches its asymptotic convergence rate. \\}
\note{In our experiments, we find that the regression-based procedure is very poor when $\Gamma$ is small. \\}
\note{It means that when we don't have enough data points, the regression-based procedure is particularly bad. \\}
\note{As we increase the total simulation budget $\Gamma$, the regression-based procedure improves fast, and eventually matches the asymptotic convergence rate. \\}

@@

\begin{frame}{Total Computation Time}

    \begin{figure}
        \includegraphics[width=\textwidth]{../project1/figures/figure11b.png}
        \caption{Total computation time for different procedures}
    \end{figure}

    \note{This figure shows the total computation time for different procedures. \\}
    \note{It includes the time for both the simulation, training, validation, and everything else. \\}
    \note{The total computation time of the standard procedure does not increase fast with $\Gamma$. \\}
    \note{It is because the computation time of estimating the optimal $M$ and $N$ does not increase for larger $\Gamma$. \\}
    \note{The regression procedure and kernel smoothing procedure are the most efficient among metamodel-based procedures. \\}
    \note{Our implementation of the kernel smoothing procedure is efficient because we only run the cross-validation once for all macro replications. \\}
    \note{The likelihood ratio procedure and KRR procedure are more expensive due to their additional costs}
    \note{The likelihood ratio weight calculations increase quadratically with $\Gamma$. \\}
    \note{For KRR, the cost of cross-validation is so high that running it once still takes a very long time. \\}
    \note{In the thesis includes a detailed analysis of the computational complexity of different procedures in terms of $\Gamma$.}

@@

@@

    \note{We first consider the traditional regression metamodels. \\}
    \note{On the left is the QQ plot between the true loss labels and the predicted loss labels for MLR, and on the right is the QQ plot for QPR. \\}
    \note{MLR and QPR metamodels make inaccurate loss predictions. \\}
    \note{Feature engineering is hardly feasible for our 240-dimensional $X$. \\}
    \note{We will see whether deep neural networks can make more accurate loss predictions. \\}

@@

    \note{This table summarizes the results of our experiments. \\}
    \note{It is the most interesting finding that DNN metamodels with suitable architectures can \textbf{cut through the noise} in training labels. \\}
    \note{The average MSEs of metamodels on the training, test, and true data are reported in the three columns, respectively. \\}
    \note{Note that the training and test labels are noisy. The simulation noise is what makes them noisy, which is less present in the true data. \\}
    \note{First, we note that DNN metamodels has lower training errors than the traditional regression metamodels. \\}
    \note{This is expected because DNNs are more flexible and expressive than traditional regression metamodels. \\}
    \note{What is more interesting is the comparison between the test and true errors. \\}
    \note{We find that the true errors of RNN-based metamodels are lower than their test errors. \\}
    \note{This suggests that the RNN and LSTM metamodels are able to cut through the simulation noise and learn the true feature-label relationship. \\}
    \note{When we take another look at the table, we find that for LSTM, its true error is even lower than its training error. \\}
    \note{This is another evidence that LSTM metamodels are able to cut through the noise in training labels and fit to the true relationship. \\}
    \note{We also find that the LSTM metamodel is able to cut through the noise in training labels and fit to the true relationship. \\}
    \note{We have another finding for RNN. RNN metamodels suffer from the vanishing gradient problem. \\}
    \note{This is why we make separate rows for well-trained RNNs and poorly-trained RNNs. \\}
    \note{We will discuss this in the next slide. \\}

@@

    \note{We now consider the tail scenario identification for the two-stage procedure. \\}
    \note{On the left is the tail scenario identification for the traditional regression metamodel, and on the right is the tail scenario identification for the LSTM metamodel. \\}
    \note{We compare the two-stage procedure with the standard procedure. \\}
    \note{When colored lines cross the grey dashed line, it means that the metamodel is able to identify the tail scenarios as well as the standard procedure. \\}
    \note{We find that the LSTM metamodel is able to surpass the standard procedure on tail identification with a less than $5\%$ safety margin. \\}
    \note{However, the traditional regression metamodel is unable to identify the tail scenarios even with a high safety margin. \\}
    \note{A well-trained RNN is better than a FNN, but the poorly-trained RNN is far worse.}
    \note{This further motivates the use of LSTM metamodels over RNN metamodels. \\}

@@

    \note{We now consider the sensitivity testing for the DNN metamodels. \\}
    \note{We are interested in the sensitivity of the DNN metamodels to the noise level in training labels and the number of training samples. \\}
    \note{In a simulation experiment, we can vary them to see how the DNN metamodels perform. \\}
    \note{We consider the noise level in training labels by varying the number of inner replications $N'$. \\}
    \note{We consider the number of training samples by varying the number of outer scenarios $M$. \\}
    \note{We consider two LSTM metamodels of different capacities because we want to see if the capacity of the metamodel matters for overfitting. \\}
    \note{The high-capacity LSTM metamodel has 10 times more parameters than the regular LSTM metamodel}

@@

    \textbf{Common Techniques:}
    \begin{itemize}
        \item   \textbf{Fine-tuning:} a model pre-trained on a source task is used as a starting point for a target task.
        \item   \textbf{Layer freezing:} only part of the model is fine-tuned.
        \item   \textbf{Multi-task learning:} perform training on multiple tasks simultaneously.
    \end{itemize}

>>>>

\jointwork{Supervised by Prof. Tony Wirjanto and Prof. Mingbin Feng}
\conference{Thesis Defense \\ University of Waterloo \\ 2025}

@@

% Title page, navigation surpressed, no page number
{
\beamertemplatenavigationsymbolsempty
\begin{frame}[plain]
\titlepage

\note{Thank you for attending my thesis defense. \\}
\note{Today, I will present my work on resilient machine learning approaches for fast risk evaluation and management of financial portfolios and variable annuities. \\}
\note{I am supervised by Prof. Tony Wirjanto and Prof. Mingbin Feng. \\}

\end{frame}
}

{
\beamertemplatenavigationsymbolsempty
\defbeamertemplate*{headline}{miniframes theme no subsection no content}
{ \begin{beamercolorbox}{section in head/foot}
    \vskip\headheight
  \end{beamercolorbox}}
\begin{frame}{Outline}
\tableofcontents

\note{My thesis focuses on improving nested simulation with supervised learning metamodels. It consists of three parts. \\}
\note{In the first part, I focus on comparing different nested simulation procedures for risk evaluation of financial derivatives. \\}
\note{In the second part, I examine deep neural network based nested simulation procedures for risk management of variable annuities. \\}
\note{In the third part, I propose a transfer learning based nested simulation procedure for fast adaptation to new variable annuity contracts.}
\end{frame}
}
\addtocounter{framenumber}{-2}

@@

Involves two levels of Monte Carlo simulations:
\begin{itemize}
    \item Outer: underlying risk factors, $X_i \sim F_X$
    \item Inner: scenario-wise losses, $Y_{ij} \sim F_{Y|X_i}$
\end{itemize}

With an expensive total simulation budget $\Gamma = M \cdot N$:
$$\hat{L}_{N, i} = \frac{1}{N} \sum_{j=1}^N Y_{ij}; ~~~ Y_{ij} \sim F_{Y|X_i} $$

\begin{itemize}
    \item Uses inner sample mean to estimate $L(X_i)$.
\end{itemize}

\note{In the context of quantitative risk management, nested simulation is used for risk evaluation of financial derivatives and insurance products. \\}
\note{The nested structure is due to the loss random variable being a conditional expectation of the portfolio loss given the underlying risk factors. \\}
\note{In our problems, the loss $L$ depends on the risk factors $X$ and is the conditional expectation of the portfolio loss given the risk factors. \\}
\note{More specifically, $X$ is $d$-dimensional and $L$ is 1-dimensional. \\}
\note{Estimating a risk measure, $\rho(L)$, involves two levels of Monte Carlo simulations: a outer level generates underlying risk factors, and a inner level generates scenario-wise samples of portfolio losses. \\}
\note{The standard nested simulation procedure uses the inner sample mean under that scenarioto estimate $L(X_i)$. \\}
\note{The nested structure makes the estimation computationally expensive. \\}
\note{In this thesis, we focus on examining and improving the efficiency of nested simulation procedures mostly by pooling with supervised learning metamodels.}

@@

% \begin{frame}{Common Risk Measures}

% \begin{itemize}
%     \item Smooth $h$, e.g., quadratic tracking error,

%     $$ \rho(L) = \mathbb{E} \left[ (L - b)^2 \right]; $$

%     \item hockey-stick $h$: mean excess loss,

%     $$ \rho(L) = \mathbb{E} \left[ L \cdot \mathbbm{1}_{\{L \geq u\}} \right]; $$

%     \item indicator $h$: probability of large loss,

%     $$ \rho(L) = \mathbb{E} \left[ \mathbbm{1}_{\{L \geq u\}} \right]; $$

%     \item Value at Risk (VaR),

%     $$ \rho_\alpha(L) = Q_\alpha(L) = \inf \{ u: \mathbb{P}(L \leq u) \geq \alpha \}; $$

%     \item Conditional Value at Risk (CVaR) \footnotemark,

%     $$ \rho_\alpha(L) = \mathbb{E} \left[ L \vert L \geq Q_\alpha(L) \right]. $$

%     \end{itemize}

%     \footnotetext{Note: If $Q_\alpha(L)$ falls in a probability mass, $\rho(L) = \frac{(\beta - \alpha)Q_\alpha(L) + (1- \beta) \mathbb{E} \left[ L \vert L \geq Q_\alpha(L) \right]}{1-\alpha}$.}

% \note{We consider different types of risk measures in this thesis. \\}
% \note{The first three are in expectation form, and the last two are quantile-based risk measures. \\}
% \note{The difference between the first three is the choice of the function $h$. \\}
% \note{$h$ can be a smooth function, a hockey-stick function, or an indicator function. \\}
% \note{The examples of risk measures are quadratic tracking error, mean excess loss, and probability of large loss. \\}
% \note{The last two are value at risk and conditional value at risk, which are widely used in practice. \\}
% \note{Note that the definition of CVaR requires the knowledge of the Value at Risk.}

@@

% \begin{frame}{Standard Nested Simulation}

% $$\hat{L}_{N, i} = \frac{1}{N} \sum_{j=1}^N Y_{ij}; ~~~ Y_{ij} \sim F_{Y|X_i} $$

% \begin{itemize}
%     \item Uses inner sample mean to estimate $L(X_i)$.
%     \item Proposed by \citet{gordy2010nested}; finds optimal growth order of $M$ and $N$.
%     \item \citet{zhang2021bootstrap} estimate the optimal $M$ and $N$ using a bootstrap method.
%     \item Computationally expensive and potentially \textbf{wasteful} use of budget.
% \end{itemize}

% \note{In our first project, we focus on comparing existing nested simulation procedures for different types of risk measures. \\}
% \note{They differ in the way they estimate $L(X_i)$, the portfolio loss conditional on the risk factors $X_i$. \\}
% \note{The standard nested simulation procedure uses the inner sample mean under that scenarioto estimate $L(X_i)$. \\}
% \note{The optimal growth order of $M$ and $N$ is found by \citet{gordy2010nested}. \\}
% \note{For a growing total simulation budget $\Gamma$, $M$ should grow in the order of $\Gamma^{2/3}$, and $N$ should grow in the order of $\Gamma^{1/3}$. \\}
% \note{For a finite-sample case, the bootstrap method is used to estimate the optimal $M$ and $N$ by \citet{zhang2021bootstrap}. \\}
% \note{Since the standard procedure only uses the information from the same scenario, it is potentially \textbf{wasteful} use of budget. \\}
% \note{Subsequent works focus on improving the efficiency of nested simulation by pooling inner replications from different outer scenarios.}

@@

\note{Pooling can happen in different ways, and the key idea is to use the information from different outer scenarios to improve the efficiency of nested simulation. \\}
\note{A supervised learning based nested simulation procedure pools with these steps. \\}
\note{First, the user runs the standard nested simulation procedure to generate a set of outer scenarios and inner replications. \\}
\note{Then, the outer scenarios and inner sample means can be treated as features and labels for metamodel training. \\}
\note{Finally, instead of the inner sample mean from the standard procedure, the metamodel predictions estimate the portfolio loss and the risk measures. \\}
\note{Most existing works focus on the convergence rate of their estimators in terms of the simulation budget. \\}
\note{However, we find that there is a cost of pooling that comes from metamodel training, validation, and prediction. }

@@

\note{We start by comparing the asymptotic rates of different nested simulation procedures. \\}
\note{These are the convergence rates of the MSE of the risk measure estimator when the total simulation budget $\Gamma$ goes to infinity. \\}
\note{The standard procedure has the same convergence rate of $\mathcal{O}(\Gamma^{-2/3})$ for all types of $h$ considered. The rates in red are not available in the original works but are derived in this thesis. \\}
\note{The regression-based procedure has $\mathcal{O}(\Gamma^{-1})$ convergence rate for smooth and hockey-stick $h$. \\}
\note{It is achieved by only allowing $M$ to grow. In their experiments, $N$ is set to equal a constant. \\}
\note{The kernel smoothing procedure is the only procedure that depends on the asset dimension $d$. \\}
\note{When risk factor $X$ is high-dimensional, the kernel smoothing procedure may converge slower than the standard procedure. \\}
\note{The kernel ridge regression procedure and likelihood ratio procedure have the same convergence rate as the regression-based procedure. \\}
\note{However, in our numerical experiments, we find that their additional cost of pooling may offset their benefit of fast asymptotic rates. \\}
\note{Additionally for KRR, the authors analyze convergence of absolute error in probabilistic order instead of MSE.}

@@

\note{In most theoretical works, the focus is on the convergence rate of the MSE of the risk measure estimator. \\}
\note{However, for KRR, the authors analyze convergence of absolute error in probabilistic order. \\}
\note{One of the contributions of this thesis is to bridge the gap between MSE and absolute error convergence. \\}
\note{We first define the convergence in MSE and the convergence in probabilistic order. \\}
\note{And we show that the convergence rate of $\Gamma^{-\xi}$ in MSE of the risk measure estimator implies the convergence rate of $\Gamma^{-\frac{\xi}{2}}$ in probabilistic order. \\}
\note{We also show that the converse is not necessarily true.}

@@

% \begin{frame}{Experiment Design}

% We compare 5 nested simulation procedures 
% \begin{itemize}
%     \item Standard nested simulation
%     \item Regression-based
%     \item Kernel smoothing
%     \item Likelihood ratio
%     \item Kernel ridge regression
% \end{itemize}

% And their empirical convergence across different:
% \begin{itemize}
%     \item Risk measures
%     \item Option types
%     \item Asset dimensions
%     \item Asset models (GBM vs. Heston)
%     \item Regression bases (only for the regression-based procedure)
% \end{itemize}

% \note{Previously, we worked on the asymptotic convergence rates of different nested simulation procedures. \\}
% \note{In practice, the total simulation budget is finite.}
% \note{When the budget is finite, the theoretical convergence rate may not be the only thing we should consider. \\}
% \note{We should also consider the cost of pooling itself. \\}
% \note{The additional cost of pooling inner replications is not considered in the theoretical analysis of the previous works. \\}
% \note{Neither it is mentioned in their numerical experiments. \\}
% \note{More specifically, we have found that different research uses different numerical experiments to show the empirical rate of different procedures. \\}
% \note{We aim to provide a fair and comprehensive analysis of the finite-sample performance of different nested simulation procedures. \\}
% \note{Our numerical experiments test the empirical convergence rates of different nested simulation procedures across different risk measures, option types, asset dimensions, and asset models.}

\begin{figure}
    \includegraphics[width=0.9\textwidth]{../project1/figures/figure2a.png}
\end{figure}

\note{Previously, we worked on the asymptotic convergence rates of different nested simulation procedures. \\}
\note{In practice, the total simulation budget is finite.}
\note{When the budget is finite, the theoretical convergence rate may not be the only thing we should consider. \\}
\note{This figure shows the performance of all 5 procedures in their estimation of the quadratic tracking error for the portfolio of European call options.}
\note{It is a log-log plot of the MSE of the estimator versus the total simulation budget $\Gamma$.}
\note{For each procedure, we run 1000 macro replications of Monte Carlo simulations with different total simulation budgets. \\}
\note{For each data point on the plot, 1000 independent replications of the procedure are used to estimate the MSE. \\}
\note{The slopes of the lines estimates the empirical convergence rates of the procedures. \\}
\note{In the legend records the slopes of the lines and the corresponding asymptotic convergence rates of the procedures. \\}
\note{In the most basic case, most procedures have their empirical convergence rates match their asymptotic convergence rates. \\}
\note{However, as we will see after, this is not the case for all experiments. }

@@

    \begin{figure}
        \includegraphics[width=0.9\textwidth]{../project1/figures/figure2b.png}
    \end{figure}

    \begin{itemize}
        \item Regression and kernel converge faster than their asymptotic rates
        \item Results of other experiments are in the thesis (Figure 2.3 - Figure 2.11)
    \end{itemize}

\note{This figure shows the same experiment as the previous figure, but we further increase the asset dimension to 20. \\}
\note{The standard procedure is not affected by the asset dimension. \\}
\note{The kernel smoothing procedure actually converges faster than what its asymptotic convergence rate suggests. \\}
\note{The regression-based procedure is the only procedure that actually converges faster when the asset dimension is high. We will do a deeper dive into it in the following experiment. \\}
\note{It is also worth noting that the likelihood ratio and KRR procedures are dimension-independent. \\}
\note{While it is not shown in their empirical rate, due to their additional cost of pooling inner replications, we are not able to run these procedures when $\Gamma$ is large. \\}
\note{In the thesis, we provide a detailed analysis of the sensitivity for all 5 procedures with more than 500 experiments. }

@@

% \begin{frame}{Fast Convergence of Regression-based Procedure}

% \begin{figure}
%     \includegraphics[width=\textwidth]{../project1/figures/figure3.png}
% \end{figure}

% \begin{itemize}
%     \item Higher initial convergence rate
%     \item Stabilizes to match asymptotic rate at higher budgets
%     \item Consistent across different asset dimensions
% \end{itemize}

% \note{To investigate the regression-based procedure, we further increase the total simulation budget $\Gamma$. \\}
% \note{The figure shows the performance of the regression-based procedure in the estimation of the quadratic tracking error for the portfolio of European call options. \\}
% \note{We separate different values of $\Gamma$ into two different parts. \\}
% \note{The first part is when $\Gamma$ is small. \\}
% \note{The second part is when $\Gamma$ is large. \\}
% \note{The first part is when the regression-based procedure has a higher convergence rate than its asymptotic convergence rate. \\}
% \note{The second part is when the regression-based procedure stabilizes and matches its asymptotic convergence rate. \\}
% \note{In our experiments, we find that the regression-based procedure is very poor when $\Gamma$ is small. \\}
% \note{It means that when we don't have enough data points, the regression-based procedure is particularly bad. \\}
% \note{As we increase the total simulation budget $\Gamma$, the regression-based procedure improves fast, and eventually matches the asymptotic convergence rate. \\}

@@

@@

\begin{frame}{Additional Computational Costs}

    \begin{figure}
        \includegraphics[width=\textwidth]{./time_comparison.png}
        \caption{Additional computation time for different procedures}
    \end{figure}

    \note{This figure shows the additional computation time for different procedures. \\}
    \note{It includes the time for training, validation, and everything else. \\}
    \note{The total computation time of the standard procedure does not increase fast with $\Gamma$. \\}
    \note{It is because the computation time of estimating the optimal $M$ and $N$ does not increase for larger $\Gamma$. \\}
    \note{Regression is the most efficient among metamodel-based procedures. \\}
    \note{The likelihood ratio and KRR are more expensive due to their additional costs. \\}
    \note{The likelihood ratio weight calculations increase quadratically with $\Gamma$. \\}
    \note{For KRR, the cost of cross-validation is too high. \\}
    \note{In the thesis includes a detailed analysis of the computational complexity of different procedures in terms of $\Gamma$.}

@@

    \note{We first take a look at their performance with QQ plots. \\}
    \note{On the left is the QQ plot between the true loss labels and the predicted loss labels for QPR, and on the right is the QQ plot for LSTM. \\}
    \note{Feature engineering for QPR is hardly feasible for our 240-dimensional $X$. Its poor performance is expected. \\}
    \note{DNN metamodels are more flexible, and they can make more accurate loss predictions. \\}

@@

@@

    \note{This table summarizes the results of our experiments. \\}
    \note{It is the most interesting finding that DNN metamodels with suitable architectures can \textbf{cut through the noise} in training labels. \\}
    \note{The average MSEs of metamodels on the training, test, and true data are reported in the three columns, respectively. \\}
    \note{Note that the training and test labels are noisy. The simulation noise is what makes them noisy, which is less present in the true data. \\}
    \note{First, we note that DNN metamodels has lower training errors than the traditional regression metamodels. \\}
    \note{This is expected because DNNs are more flexible and expressive than traditional regression metamodels. \\}
    \note{What is more interesting is the comparison between the test and true errors. \\}
    \note{We find that the true errors of RNN-based metamodels are lower than their test errors. \\}
    \note{This suggests that the RNN and LSTM metamodels can learn the true feature-label relationship. \\}
    \note{When we take another look at the table, we find that for LSTM, its true error is even lower than its training error. \\}
    \note{This is another evidence that LSTM metamodels are able to cut through the noise in training labels and fit to the true relationship. \\}
    \note{RNN metamodels suffer from the vanishing gradient problem. This is why we make separate rows for well-trained RNNs and poorly-trained RNNs. }

@@

    \note{We now consider the tail scenario identification and the CVaR estimation for the two-stage procedure. \\}
    \note{On the left is the tail scenario identification plot, and on the right is the plot of CVaR. \\}
    \note{We compare the two-stage procedure with the standard procedure. \\}
    \note{When colored lines cross the grey dashed line, it means that the metamodel is able to identify the tail scenarios as well as the standard procedure. \\}
    \note{And it should be able to estimate the CVaR as accurately as the standard procedure. \\}
    \note{We find that the LSTM metamodel is able to surpass the standard procedure on tail identification with a less than $5\%$ safety margin. \\}
    \note{However, the traditional regression metamodel is unable to identify the tail scenarios even with a high safety margin. \\}
    \note{A well-trained RNN is better than a FNN, but the poorly-trained RNN is far worse.}
    \note{The CVaR estimation plot shows similar results.}

@@

    \note{We now dig deeper into the research questions asked: how well do deep learning metamodels learn from noisy data? \\}
    \note{We study this phenomenon by considering the sensitivity testing for the DNN metamodels. \\}
    \note{The sensitivity of the DNN metamodels to the noise level in training labels and the number of training samples. \\}
    \note{In a simulation experiment, we can vary them to see how the DNN metamodels perform. \\}
    \note{We control the noise level in training labels by varying the number of inner replications $N'$. \\}
    \note{We control the number of training samples by varying the number of outer scenarios $M$. \\}
    \note{We consider two LSTM metamodels of different capacities because we want to see if the capacity of the metamodel matters for overfitting. \\}
    \note{The high-capacity LSTM metamodel has 10 times more parameters than the regular LSTM metamodel}

@@

    \textbf{Common Techniques:}
    \begin{itemize}
        \item   \textbf{Fine-tuning:} a model pre-trained on a source task is used as a starting point for a target task.
        \item   \textbf{Layer freezing:} only part of the model is fine-tuned.
    \end{itemize}
